\chapter*{Acknowledgements \label{sec:acknowledgements}}

Now that my time as a PhD student is coming to an end, it is time to reflect for a moment on the
past 3.5 years. So much happened, so many people crossed my path. It all contributed to shaping the
person I am today. I like to think it made me a better person. It definitely made me a more
knowledgeable person. 
I learned so many new things, from physics and organization to just remembering to take the time to
have fun every once in a while. 
Before the bravest among you embark on the effort to read through this thesis, I would like to give
my thanks to everyone that helped me in any way throughout these past years. 

First of all, I would like to thank my promotors Dirk and Michael, for allowing me to plunge into
the adventure that is a PhD. Your doors were always open to answer my questions, or to give me
advice. And above all, thank you for sending me off to CERN! That year-and-a-half was an
experience I will never forget. 

Sezen, without you this thesis could not have happened. I cannot thank you enough for adopting me
as your student, and convincing me to join you in the razor boost effort. Working with you is
always a pleasure, and I hope we can continue working on some joint projects in the future. Since
Lukas introduced us way back in 2010, we have become great friends, and that is a gift to be
treasured. 

Harrison, every time we chat I feel like I have learned something. Your colourful way of expressing
yourself never fails to bring a smile to my face. I am very honoured to have worked with you on
this analysis, pioneering the way towards a proper statistical treatment. 

I would also like to thank Maurizio, and the rest of the razor team, for providing help and insight
whenever we were stuck. 

Of course, a big thank you goes to all my friends. To the people of the INW, thank you for making
it a fun workplace. To my friends at CERN, thank you for the crazy evenings at Meyrinoise, and the
always entertaining Penn dinners. 
Bedankt aan al mijn krachtbalvrienden, de trainingen en wedstrijden die we samen meegemaakt hebben
waren de ideale ontspanning tussen de drukte door. 
Karen en Mieke, met al onze drukke schema's hebben we elkaar niet zo vaak kunnen zien, maar ik wil
dat jullie weten dat ik onze vriendschap koester. Ik hoop dat we in de toekomst ook in contact
kunnen blijven!


Een dikke merci ook aan mijn ouders en broer. Jullie zorgden ervoor dat ik het contact met de
`echte wereld' niet verloor. 
Dankjewel voor alle steun doorheen de jaren. Zonder jullie had ik dit niet gekund! 


Brett, since I met you, my life has never been the same. We have seen and done so many fun things
over the past couple of years. 
Thank you for being there through all the mad times, rushing to meet one or the other deadline.
The last couple months were not easy, with us living 800\unit{km} apart. The evenings spent talking
until past midnight about all our future plans really kept me going. 
Now, we are embarking on a brand new and super exciting adventure in Chicago. 
We cannot know what the future will hold, but I know that we can conquer anything together!


