%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Explanation of control regions
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{\texorpdfstring{$T$}{T} region \label{sec:boost_T_region}}

The $W$ tagger defined above will be used to select events in the signal region and in the TTJets
control region (see section~\ref{sec:Sregion} and \ref{sec:Tregion}). In both of those regions, we
expect to have real hadronically decaying $\W$ bosons, mainly coming from the $t\bar{t}+$jets
background. 

\subsection{\texorpdfstring{$Q$}{Q} region}

Another background in the signal region, is QCD multijet production, for which we have defined a
dedicated control region (section~\ref{sec:Qregion}). Multijet events do not have real $\W$ bosons;
any CA8 jet that passes the W tagger, is thus by definition a misidentified or \textit{fake} $\W$. 
For the QCD control region we define ``anti-tagged $W$'s'' ($aW$), in which we invert the
n-subjettiness cut but leave the other cuts the same. 
As we do not expect QCD events to have the two-prong jet-substructure of actual boosted $W$'s, this
will enhance the statistical power of the control region. 

\subsection{\texorpdfstring{$W$}{W} region}

Our final background in the signal region for which we have a dedicated control region, is
$\W(\rightarrow l\nu)$+jets. 
As the $\W$ boson decays leptonically, any events passing the signal selection must contain a CA8
jet that is misidentified as a boosted $\W$. 
We do not expect hadronically decaying $\W$ bosons to be a large background, as those events would
usually not come with $\cPqb$-jets, and would have very small missing energy. Even though we do not
explicitely require a minimal \ETm, our $R^2$ cut is highly correlated with \ETm, and induces
a minimal cut of around 100\GeV, as can be seen in figure~\ref{fig:DataMC_SignalRegion_mdphig0p5}.  
We have verified this assumption by checking the contribution of $\W(\rightarrow
q\bar{q}')+b\bar{b}$ using simulation, and found that no events pass our signal selection. 
For the $W$+jets control region, which is defined by requiring the presence of a lepton, we only
keep the mass window requirement to select CA8 jets that fake a boosted $\W$ boson. We will denote
these mass-tagged $\W$'s as $Y$. We do not use any cut on the n-subjettiness variables to increase
the number of events available in the control region. 
